---
title: "Text Mining Project"
subtitle: "Sentiment Analysis of Tweets About the Tesla Cybertruck"
author: "Mark Riley"
date: "12/8/2019"
output: 
  word_document:
    toc: yes
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary libraries
library(tidyverse)
library(wordcloud)
library(tm)
library(syuzhet)
library(ggplot2)
library(rtweet)
library(maps)
library(lubridate)
library(scales)
library(gridExtra)

# Go Eagles!
laxMaroon = "#830019"
laxGray = "#969799"

setwd("~/Dropbox/UWLax/DS745/Text Mining Project")

# Load the tweets
allTweets <- read_twitter_csv("CybertruckTweets.csv", unflatten = TRUE)

# Convert UTC to EDT
allTweets %<>% dplyr::mutate(created_at = as_datetime(created_at, tz = "UTC")) %>% 
      dplyr::mutate(created_at = with_tz(created_at, tzone = "America/New_York"))

# Produce lat and lng coordinates
allTweets <- lat_lng(allTweets)

# Save tweets to another data set "text"
text = allTweets$text

# Remove names
text = gsub("(RT|via)((?:\\b\\W*@\\w+)+)","",text)

# Remove html links
text = gsub("http[^[:blank:]]+","",text)

# Remove people names
text = gsub("@\\w+","",text)                              

# Remove punctuations
text = gsub("[[:punct:]]","",text)

# Remove whitespace
text = trimws(text, which = c("both", "left", "right"))

# Remove digits
text = gsub('[[:digit:]]+', '', text)

# Remove line breaks
text = gsub("[\r\n]", "", text)

# Remove not readable standard text
text = iconv(text, to = "ASCII//TRANSLIT")

# Remove not readable standard text
text = iconv(text, "ASCII", "UTF-8", sub="")

# Lower case
text = tolower(text)

# Put the text back into the dataframe
allTweets$text <- text

# Remove the NA values from the text
text <- text[!is.na(text)]

# Remove NA values from the dataframe
allTweets  <- allTweets[!is.na(allTweets$text),]

# Convert tweets to corpus
corpus = Corpus(VectorSource(text))

# Remove stopwords
corpus = tm_map(corpus, removeWords, stopwords("english"))

# Remove whitespace
corpus = tm_map(corpus, stripWhitespace)

# Convert corpus to Term Document matrix 
uniqwords = as.matrix(TermDocumentMatrix(corpus))

# Count frequency of words in each tweet
wordfreq = sort(rowSums(uniqwords), decreasing = TRUE)

# Word frequency dataframe
WCinput = data.frame(word = names(wordfreq), freq = wordfreq)

# Get sentiment
sent.v <- get_sentiment(as.character(text))

# Combine the tweet text, the NRC sentiment, sentiment,
# and verified account status data into a data frame
sent.df <- data.frame(text, sent.v, allTweets$created_at, allTweets$verified)
colnames(sent.df)[2:4] = c("score","created_at", "verified")

# Create a rating (positive, neutral, negative) for each tweet
sent.df <- sent.df %>% 
  mutate(rating = case_when(
    score == 0 ~ "neutral",
    score > 0 ~ "positive",
    score < 0 ~ "negative",
    TRUE ~ "error"
  ))
```

# Dataset Description  
The dataset consists of 28,756 tweets that contain the 'cybertruck' keyword. The tweets were gathered from Twitter by the authors using the rtweet library on December 2, 2019. All of the tweets are in English and geocoded to the United States of America. Retweets are excluded. The timeline of the tweets extend from approximately 4:11 PM EST on November 23, 2019 until 6:45 PM EST on December 2, 2019.  

# Problem Description  
Tesla debuted their initial entry into the pickup truck market on November 21, 2019. Tesla CEO Elon Musk unveiled the Cybertruck to a live audience in Los Angeles and broadcast via the internet to the rest of the world at 8:00 PM PST.  

The Cybertruck design is unlike any other pickup truck currently or previously offered by a manufacturer. It's futuristic, angular design was met with mixed reviews. In an interview with CNN^1^, Matt DeLorenzo, Senior Executive Editor at Kelley Blue Book said, "It will be a niche product at best and poses no threat in the pickup market as we know it today." In the same CNN interview, Chelsea Sexton, an analyst who covers the electric vehicle market, said "From a specification standpoint, I believe that's probably what they're aiming for, but no question that body style, that is not a high-volume product."

The US pickup truck market in 2018 represented over $72B in revenue, according to Forbes^2^. Tesla will have competition soon from new and existing car manufacturers. According to Reuters^3^, Ford aims to sell its electric pickup in Fall 2020, and GM will build a family of premium electric pickups the following year. Rivian, a start up electric vehicle manufacturer backed by Ford, also plans its own electric pickup next Fall. Tesla plans to deliver the first Cybertrucks in late 2021.  

Tesla and its competitors will be interested to gauge the sentiments of the public as they finalize their initial electric pickup truck models. It will be helpful to these manufacturers to determine if there is a market for such a non-traditional vehicle design like the Tesla Cybertruck. An interesting aspect is how is sentiment changing over time in the days after the Cybertruck unveiling? Is positive sentiment increasing or decreasing over time as the public gets past the initial shock of the vehicle design?  

Vehicle manufacturers may also be interested to know the sentiment for verified accounts vs. non-verified accounts. Twitter describes^4^ a verified account as, "An account may be verified if it is determined to be an account of public interest. Typically this includes accounts maintained by users in music, acting, fashion, government, politics, religion, journalism, media, sports, business, and other key interest areas." Given that verified accounts are related to news outlets, celebrities, and others with social influence, manufacturers may be interested in targeting those sources if there is a significantly more negative sentiment from verified accounts compared to non-verified accounts.  

# Data Visualization
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Cybertruck Tweet Locations"}
# Plot a map of tweet locations
par(mar = rep(.25, 4))
map("state", lwd = .25)

# Plot lat and lng points onto state map
with(allTweets, points(lng, lat, pch = 16, cex = .5, col = rgb(0.502, 0, 0.161, .2)))
```

The above figure shows where the tweets in our dataset are coming from. A stronger color opacity represents more tweets coming from that location. We can see that unsurprisingly many tweets are coming from major metropolitan areas. Of the 28,756 observations in the data, only 1,670 (`r paste0(round(1670/28756, 4)*100, "%")`) had latitude and longitude data available.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Tweet Distributions"}
# Plot the number of tweets per day
ggplot(data = allTweets, aes(x = date(created_at))) +
        geom_bar(aes(fill = ..count..), binwidth = 1) +
        theme(legend.position = "none", panel.background = element_blank()) +
        xlab("Tweet Date (EST)") + ylab("Number of Tweets") + 
        scale_fill_gradient(low = laxGray, high = laxMaroon)
```

The figure above shows the number of tweets by day. A more maroon color is related to a relatively high number of tweets for the day as compared to a lower number of tweets represented by the color gray. Keep in mind that tweets on November 23 and December 2 (earliest and latest dates in the dataset, respectively) do not represent a full 24 hours of data. The number of tweets steadily declines after the first two days of full data until stabilizing around the fifth day of full data.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Cybertruck Tweet Word Cloud"}
# Generate the word cloud
set.seed(255)
wordcloud(words = WCinput$word, freq = WCinput$freq, min.freq = 3, scale=c(4,.5),
          max.words=175, random.order = FALSE, rot.per = 0.35,
          colors = c(laxGray, laxMaroon))
```

The word cloud for the dataset shows a mix of positive and negative sentiments. Words like "want", "ordered", "future", "cool", and "great" show the positive sentiment. Negative sentiment is represented by words such as "ugly" and "bad". There are also references to "windows", "glass", and "break". During the Cybertruck reveal, Tesla had a failed demonstration of the window glass strength, breaking two windows on the driver's side of the prototype vehicle with a metal ball. Initial review of the word cloud shows that positive and neutral sentiments seem to be more frequent than negative terms.  

# Variables  
The variable we used for assessing the sentiment of each tweet was the tweet 'text' field. The field 'created_at', representing the time stamp of the tweet, was used for assessing the number of tweets and sentiment over time. The binary 'verified' field is used to determine if a Twitter account is verified.  

# Methodology  
We performed the following steps to assess the sentiments of the tweet text in our dataset.  

1. Converted the tweet time stamps from Coordinated Universal Time (UTC) to Eastern Time, America/New York.
2. Performed cleaning of the Tweet text by removing names, html links, twitter handles, punctuation, white space, digits, and line breaks.
3. Converted the text to lower case.
4. Removed any observations where text cleanup removed all of the Tweet text.
5. Calculated the sentiment of each remaining Tweet.
6. Added ratings for each Tweet based on the sentiment score.
    + Score > 0 = positive.
    + Score < 0 = negative.
    + Score = 0 = neutral.

To produce the word cloud, we:

1. Created a corpus of the tweet text.
2. Removed stop words and white space.
3. Created a term document matrix.
4. Counted the frequency of each word.
5. Generated the word cloud.

# Findings and Discussion  
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Cybertruck Tweet Sentiment Totals"}
# Plot the total number of tweets by sentiment rating
ggplot(data = sent.df, aes(x = rating)) +
        geom_bar(fill = laxMaroon, color = laxGray, stat = "count") +
        theme(legend.position = "none", panel.background = element_blank()) +
        xlab("Sentiment") + ylab("Number of Tweets")
```

The figure above shows that overall the tweets in the days following the unveiling of the Tesla Cybertruck, more tweets were positive than negative or neutral. This is good news for Tesla given the unusual design of the vehicle and the issues with the glass strength demonstration.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Cybertruck Tweet Sentiments by Day"}
# Plot the percentage of tweets each day that fall into the three rating values
sent.df %>% 
  count(rating, tdate = date(created_at)) %>% 
  group_by(tdate) %>% 
  mutate(pct = n/sum(n)) %>% 
  ggplot(aes(fill = rating, y = pct, x = tdate)) +
  geom_bar(position="fill", stat="identity") +
  theme(panel.background = element_blank()) +
  xlab("Sentiment") + 
  ylab("Proportion of Tweets") +
  scale_fill_manual(values = c("black", laxGray, laxMaroon)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position="bottom") +
  geom_text(aes(label = paste0(round(pct*100), "%")), position = position_stack(vjust = 0.5), size = 3, 
            color="white", fontface = "bold")
```

The figure above shows that positive sentiment has fluctuated slightly but remained within a seven percentage point range (46% - 53%). Neutral sentiment has remained even more steady within a four percentage point range (25% - 29%). Negative sentiment was also within a seven percentage point range (21% - 28%).  

Again we would classify this result as positive for Tesla given the unusual, futuristic design of the vehicle and failed glass strength demonstration during the unveiling event.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Cybertruck Tweet Sentiments by Twitter Verification Staus"}
# Graph the sentiment by verified status
sent.df %>% 
  count(rating, verified) %>% 
  group_by(verified) %>% 
  mutate(pct = n/sum(n)) %>% 
  mutate(status = case_when(
    verified == TRUE ~ "Verified Accounts",
    verified == FALSE ~ "Non-verified Accounts",
    TRUE ~ "Error"
  )) %>% 
  ggplot(aes(x = rating, y = pct)) +
  geom_bar(position="dodge", stat="identity", fill = laxMaroon, color = laxGray) +
  facet_wrap(~status) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position="none", panel.background = element_blank()) +
  xlab("Sentiment") + ylab("Percentage of Tweets")

```

The figure above shows that sentiment between verified and non-verified Twitter accounts is nearly identical. Both types of accounts are even at roughly 50% for positive sentiment. Verified accounts have just slightly more negative sentiment than non-verified accounts. This shows that there may be an opportunity to target verified accounts to increase their positive sentiment with the intent that will influence non-verified accounts to have more positive sentiment.  

# Relevant R Outputs  
## Sentiment Score Distribution  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Cybertruck Tweet Sentiment Score Distribution"}
p1 <- ggplot(sent.df, aes(y = score)) +
  geom_boxplot(color = laxMaroon) +
  theme(legend.position="none", panel.background = element_blank()) +
  xlab("") + ylab("Sentiment Score")

p2 <- ggplot(sent.df, aes(x = score)) +
  geom_histogram(color = laxGray, fill = laxMaroon) +
  theme(legend.position="none", panel.background = element_blank()) +
  xlab("Sentiment Score") + ylab("Count")

grid.arrange(p1, p2, ncol = 2)
```

The figures above show that the sentiment scores appear to be mildly skewed to the right toward the positive sentiment. The scores range from a maximum of `r max(sent.df$score)` on the positive side to `r min(sent.df$score)` on the negative side. The mean sentiment score is `r mean(sent.df$score)`, showing some bias toward positive sentiment. The sentiment score standard deviation is `r sd(sent.df$score)`.  

Below we can see a sampling of ten tweets from each of the sentiment categories. From these examples we are able to see some of the limitations of sentiment analysis. In the positive tweets there is a one calling the Cybertruck a "joke" but praising the Model X, Tesla's full-sized SUV. The last example of negative tweets the user is saying they really want a Cybertruck will not be able to afford it. Many of the sampled neutral tweets could reasonably be classified as positive. In general the sentiment analysis seems to be relatively accurate.  

## Top 10 Positive Tweets  
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the top 10 positive tweets
knitr::kable(sent.df %>% arrange(desc(score)) %>% head(10) %>% select(text))
```

## Top 10 Negative Tweets  
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the top 10 negative tweets and filter out some colorful language
neg.df <- sent.df %>% arrange(score) %>% head(10) %>% select(text)
neg.v <- as.vector(neg.df$text)
neg.v <- gsub("fuck", "frig", neg.v)
neg.v <- gsub("shit", "poop", neg.v)
neg.v <- gsub("ass", "butt", neg.v)
neg.df <- data.frame(neg.v)
knitr::kable(neg.df)
```

## Sample of 10 Neutral Tweets  
```{r echo=FALSE, message=FALSE, warning=FALSE}
# randomly select 10 neutral tweets
set.seed(255)
knitr::kable(sent.df %>% filter(score == 0) %>% sample_n(10) %>% select(text))
```

# References
1: Valdes-Dapena, P. (2019). Tesla reveals Cybertruck, but breaks its 'unbreakable' windows during unveiling. [online] CNN. Available at: https://www.cnn.com/2019/11/22/cars/tesla-cybertruck-electric-pickup-truck/index.html [Accessed 5 Dec. 2019].

2: Forbes.com. (2019). Cybertruck Could Bring In Over $4.5 Billion In Revenue For Tesla By 2025. [online] Available at: https://www.forbes.com/sites/greatspeculations/2019/11/28/cybertruck-could-bring-in-over-45-billion-in-revenue-for-tesla-by-2025/#72894b67e560 [Accessed 5 Dec. 2019].

3: Klayman, B. (2019). Tesla picks a fight with Detroit by going after the pickup market. [online] Reuters. Available at: https://www.reuters.com/article/us-tesla-pickup/tesla-picks-a-fight-with-detroit-by-going-after-the-pickup-market-idUSKBN1XV1CN [Accessed 5 Dec. 2019].

4: Help.twitter.com. (2019). About verified accounts. [online] Available at: https://help.twitter.com/en/managing-your-account/about-twitter-verified-accounts [Accessed 5 Dec. 2019].